\section{Additional Software}
These packages aren't included in the ET distribution. 
So, you need to download them separately. 
\subsection{CVX}
Some methods of the {\it Ellipsoidal Toolbox}, namely,
\begin{itemize}
\item {\tt distance}
\item {\tt intersect}
\item {\tt isinside}
\item {\tt contains}
\item {\tt ellintersection\_ia}
\item {\tt ellunion\_ea}
\end{itemize}
require solving semidefinite programming (SDP) problems.
We use CVX (\cite{cvxhp}) as an
interface to an external SDP solver. CVX as a more reliable toolbox
for solving SDP problems of high dimensionality. 
CVX is implemented in Matlab, effectively turning Matlab into an 
optimization modeling language. Model specifications are constructed
using common Matlab operations and functions, and standard 
Matlab code can be freely mixed with these specifications. 
This combination makes it simple to perform the calculations needed 
to form optimization problems, or to process the results
obtained from their solution. CVX distribution includes two freeware 
solvers: SeDuMi(\cite{sedumi}, \cite{sedumihp})
and SDPT3(\cite{sdpt}). The default solver used in the toolbox is SeDuMi.

\subsection{MPT}
Multi-Parametric Toolbox(\cite{mpt}) - a Matlab toolbox for multi-parametric
optimization and computational geometry. MPT is a toolbox 
that defines polytope class
used in {\it ET}.
Here we list methods required MPT.
\begin{itemize}
\item {\tt distance}
\item {\tt intersect}
\item {\tt intersection\_ia}
\item {\tt intersection\_ea}
\item {\tt isinside}
\item {\tt hyperplane2polytope}
\item {\tt polytope2hyperplane}
\end{itemize}
We need MPT for all methods operating with polytopes.


\section{Installation and Quick Start}
\begin{enumerate}
\item Go to
\newline
{\tt \url{http://code.google.com/p/ellipsoids}}
\newline
and download the {\it Ellipsoidal Toolbox}.
\item Unzip the distribution file into the directory where you would like
the toolbox to be.\todo{Create a zip archive with the latest version and upload it to the project web site}
\item Unzip CVX into cvx folder next to products folder.
\item Unzip MPT into mpt folder next to products folder.
\item Read the copyright notice.
\item In MATLAB command window change the working directory to the one where
you unzipped the toolbox and type
\verbmcodef[Installation]{mcodesnippets/s_chapter04_section01_snippet01.m}
\item At this point, the directory tree of the {\it Ellipsoidal Toolbox} is
added to the MATLAB path list. In order to save the updated path list,
in your MATLAB window menu go to {\tt File} $\rightarrow$ {\tt Set Path...} and
click {\tt Save}.
\item To get an idea of what the toolbox is about, type
\verbmcodef[Basic functionality]{mcodesnippets/s_chapter04_section01_snippet02.m}

This will produce a demo of basic {\it ET} functionality: how to create
and manipulate ellipsoids.

Type

\verbmcodef[Plot ellypsoids and hyperplanes]{mcodesnippets/s_chapter04_section01_snippet03.m}
to learn how to plot ellipsoids and hyperplanes in 2 and 3D.
\newline
For a quick tutorial on how to use the toolbox for reachability analysis
and verification, type

\verbmcodef[Tutorial for reachability analysis
and verification]{mcodesnippets/s_chapter04_section01_snippet04.m}
\end{enumerate}
